\documentclass[10pt, conference]{IEEEtran}
\IEEEoverridecommandlockouts


\usepackage[utf8]{inputenc}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}

\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{subfigure}
\usepackage{float}


% make proper ticks so that one does not need to ``something" as it looks ugly imho

\usepackage[english]{babel}
\usepackage[autostyle, english=american]{csquotes}
\MakeOuterQuote{"}

\usepackage[colorinlistoftodos]{todonotes} % cool todos

% ALWAYS NEEDS TO BE THE LAST PACKAGE DEFINED
\usepackage{cleveref} % note make \Cref and not \ref
% needed to link paragraphs to Cref
\Crefname{paragraph}{Paragraph}{Paragraphs}
\crefname{paragraph}{paragraph}{paragraphs}




% Define custom colors for listings
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

% Setup for listings package
\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}
\lstset{style=mystyle}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}

\begin{document}
\include{tools/cmds}

\title{High-Performance Parallel Solver for the Futoshiki Puzzle:\\
A Multi-Paradigm Approach using MPI, OpenMP, and Hybrid Parallelization\\
\normalsize \textit{High Performance Computing for Data Science Project 2024/2025}
}

\author{\IEEEauthorblockN{Wendelin Falschlunger}
\IEEEauthorblockA{\textit{Mat. 249562} \\
\textit{University of Trento, DISI}\\
38123 Povo TN, Italy \\
w.falschlunger@studenti.unitn.it}
\and
\IEEEauthorblockN{Lorenzo Dongili}
\IEEEauthorblockA{\textit{Mat. 247204} \\
\textit{University of Trento, DISI}\\
38123 Povo TN, Italy \\
lorenzo.dongili@studenti.unitn.it}
\and
\IEEEauthorblockN{Stefano Dal Mas}
\IEEEauthorblockA{\textit{Mat. 247201} \\
\textit{University of Trento, DISI}\\
38123 Povo TN, Italy \\
stefano.dalmas@studenti.unitn.it}
}

\maketitle

\begin{abstract}
The Futoshiki puzzle, an NP-Complete variant of the Latin Square Completion Problem, presents significant computational challenges as grid size increases. This paper presents a comprehensive high-performance computing solution employing multiple parallelization paradigms. We first implement an efficient sequential solver based on the list coloring approach proposed by Şen and Diner in \cite{Sen2024Futoshiki}, which employs constraint propagation to dramatically reduce the search space. Building upon this foundation, we develop three distinct parallel implementations: (1) an MPI distributed-memory solver employing a master-worker paradigm, (2) an OpenMP shared-memory solver using task-based parallelism, and (3) a novel hybrid MPI+OpenMP solver that combines both approaches for maximum scalability. All implementations feature a sophisticated multi-level work generation strategy that dynamically adjusts the granularity of parallelism. Our experimental evaluation on a high-performance computing cluster demonstrates that the pre-coloring optimization achieves up to 14× \wen{re-check these values!} speedup over naive backtracking, while our parallel implementations show strong scalability with speedups of up to 17.5× for MPI, 15.2× for OpenMP, and 28.3× for the hybrid approach on appropriate workloads. The results establish our multi-paradigm framework as a robust and efficient solution for complex combinatorial problems.
\end{abstract}

\begin{IEEEkeywords}
Futoshiki, Constraint Satisfaction, List Coloring, OpenMP, MPI, Hybrid Parallelization, High-Performance Computing, Task-Based Parallelism
\end{IEEEkeywords}

\input{sections/01_introduction}
\input{sections/02_related_work}
\input{sections/03_solution}
\input{sections/04_experimental_evaluation}
\input{sections/05_conclusion}
\input{sections/06_future_work}

% Bibliography
\bibliographystyle{IEEEtran}
\bibliography{bibliography}

\end{document}
